---
title: "WiscSIMSLab Session Summary"
author: "Markdown Document by B.J. Linzmeier"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
citation_package: biblatex
biblio-style: chem-acs
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
keep_tex: yes
link-citations: yes
#bibliography: SIMS Report Refs.bib
always_allow_html: yes
---

  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(readxl)
library(pillar)
library(tidyr)
library(ggplot2)
library(doBy)
library(ggfortify)
library(kableExtra)
library(ggforce)
library(dplyr)
library(ggExtra)
library(svDialogs)

#Mounts <- 7 #dlg_input()$res
#Mounts <- as.numeric(Mounts)

source('GeneralSIMSImporter.R')

ExampleFilename <- "/Users/macrostrat/Desktop/SIMSdatafiles/20210224_d18O_Helser-Linzmeier/20210224_d18O_Helser-Linzmeier_final.xlsx" #file.choose() #"C:/Users/Ben/Desktop/20191021_d18O_Helser.xlsx" #file.choose()
#ExampPlugs <- readline(prompt="Enter plugs: ")
filename <- basename(ExampleFilename)

FileInput <- GeneralSIMSImporter(ExampleFilename, PlugNum = 2)

NoBlanks <- FileInput[is.na(FileInput$File)==FALSE,]
TotalAnalyses <- length(NoBlanks$File)

#### This could probably be simplified using 
#### start and end times from NoBlanks$Time
#### Maybe fix it later?
#### 

StartDate <- NoBlanks$DATETIME[1]
EndDate <- NoBlanks$DATETIME[nrow(NoBlanks)]
StartTime <- NoBlanks$DATETIME[1]
EndTime <- NoBlanks$DATETIME[nrow(NoBlanks)]


MassCalibTime <- seq(from = min(NoBlanks$DATETIME, na.rm=TRUE), to = max(NoBlanks$DATETIME, na.rm=TRUE), by = "12 hour")

#### Calculate the total analysis time####
#### based on sum of analysis lengths of less than 15 minutes

AnalysisTime <- paste(round(sum(NoBlanks$AnalysisLength[NoBlanks$AnalysisLength<15], na.rm=TRUE)/60, digits = 2), "hours")

#### Calculate the average analysis length####
#### Based on analysis lengths <15 minutes to 
#### avoid bias due to longer periods of instrument
#### shutdown (e.g. overnight/morning)

AverageAnalysis <- paste(round(mean(NoBlanks$AnalysisLength[NoBlanks$AnalysisLength<15], na.rm=TRUE), digits = 2), "min")

#### Calculate the total session time####
#### from first to last analysis without subtracting
#### time periods where no analysis is taking place

SessionTime <- round(difftime(EndTime, StartTime, units = "hours"), digits = 2)

Brackets <- FileInput[FileInput$Comment=="bracket average and 2SD"&is.na(FileInput$Comment)==FALSE,]

Samples <- FileInput[FileInput$MATERIAL=='Sample'&is.na(FileInput$File)==FALSE,]

Standards <- FileInput[FileInput$MATERIAL=='STD'&is.na(FileInput$File)==FALSE,]

StandardNum <- nrow(Standards)
SampleNum <- nrow(Samples)
BracketNum <- nrow(Brackets)
BracketAve <- mean(Brackets$SE2int)
BeamRange <- range(FileInput$IPnA, na.rm = TRUE)
BeamAverage <- mean(FileInput$IPnA, na.rm = TRUE)
Beam2SD <- 2*sd(FileInput$IPnA, na.rm = TRUE)

#### Time summary table ####
#### Combine these into a readable table

TimeSummary <- data.frame(StartTime, EndTime, AnalysisTime, SessionTime, AverageAnalysis)

#### Primary beam summary table ####
#### Combine these into a readable table

BeamSummary <- round(data.frame(BeamAverage, BeamRange[1], BeamRange[2], Beam2SD, AnalysistoAnalyis = mean(diff(FileInput$IPnA), na.rm = TRUE)), digits = 5)

####Create GGPLOT objects for important variables ####
####

Standard2SDHist <- ggplot(data = Brackets, aes(x=`SE2int`))+
  geom_histogram(binwidth = 0.05, colour="black")+
  xlab("Bracket 2SD (‰)")+
  xlim(0,1)+
  theme_bw()

StandardMeasHist <- ggplot(data = Brackets, aes(x= d18Omeas))+
  geom_histogram(binwidth = 0.05, alpha = 0.25)+
  xlab("Bracket Measured \u03B418O")+
  xlim(min(Brackets$d18Omeas)-.5, max(Brackets$d18Omeas)+.5)+
  theme_bw()
  
  

StandardMeasVio <- ggplot(data = Brackets[!is.na(Brackets$`d18Omeas`),], aes(x = GUESS.SAMP, y = d18Omeas, fill = GUESS.SAMP))+
  geom_violin(trim=FALSE)+
  geom_dotplot(binaxis='y', stackdir='center', dotsize=.5, alpha = 0.5)+
  geom_point(alpha = 0)+
  xlab("Estimated mount")+
  coord_flip()+
  ylab("Bracket Measured \u03B418O")+
  ylim(min(Brackets$d18Omeas)-.5, max(Brackets$d18Omeas)+.5)+
  theme_bw()+
  theme(legend.position = "none")
  
StandardMeasVio <- ggMarginal(StandardMeasVio, type = "histogram", margins = "x", size = 2, fill='grey', color='black', xparams = list(binwidth = 0.05), alpha = 0.25)


PrimaryBeamHist <- ggplot(data = NoBlanks, aes(x=`IPnA`))+
  geom_histogram(binwidth = 0.01, colour="black", fill="white")+
  xlab("IP (nA)")+
  theme_bw()
  
RelHydd18O <- ggplot(data = NoBlanks, aes(x = REL_Hyd, y = STDiso, color = GUESS.SAMP))+
  geom_point()+
  ylab("STD corrected \u03B418O (\u2030, VSMOW)")+
  xlab("Hyd-STD.Hyd")+
  theme_bw()+
  theme(legend.position="bottom")
  
RelYieldHyd <- ggplot(data = NoBlanks, aes(x = REL_YIELD, y = REL_Hyd, color = GUESS.SAMP))+
  geom_point()+
  xlab("Yield Relative to STD")+
  ylab("Hyd-STD.Hyd")+
  theme_bw()+
  theme(legend.position="bottom")
  
RelYieldd18O <- ggplot(data = NoBlanks, aes(x = REL_YIELD, y = STDiso, color = GUESS.SAMP))+
  geom_point()+
  ylab("STD corrected \u03B418O (\u2030, VSMOW)")+
  xlab("Yield Relative to STD")+
  theme_bw()+
  theme(legend.position="bottom")

PrimaryBeamTime <- ggplot(data = NoBlanks, aes(x = `DATETIME`, y=`IPnA`))+
  geom_line()+
  geom_point(aes(color = GUESS.SAMP))+
  geom_vline(xintercept = MassCalibTime, colour = "red")+
  scale_x_datetime()+
  ylab("IP (nA)")+
  theme_bw()+
  theme(legend.position="bottom")
  
d18OTime <- ggplot(data = NoBlanks, aes(x = `DATETIME`, y=`d18Omeas`))+
  geom_point(aes(color = MATERIAL))+
  geom_smooth(data = NoBlanks[NoBlanks$MATERIAL=='STD',], method = 'loess')+
  geom_vline(xintercept = MassCalibTime, colour = "red")+
  scale_x_datetime()+
  ylab("\u03B418O Measured (\u2030)")+
  theme_bw()+
  theme(legend.position="bottom")
  
YieldTime <- ggplot(data = NoBlanks, aes(x = `DATETIME`, y=`Yield`))+
  geom_point(aes(color = MATERIAL))+
  geom_smooth(data = NoBlanks[NoBlanks$MATERIAL=='STD',], method = 'loess')+
  geom_vline(xintercept = MassCalibTime, colour = "red")+
  scale_x_datetime()+
  ylab("Yield (Gcps/nA)")+
  theme_bw()+
  theme(legend.position="bottom")
  
SEintTime <- ggplot(data = NoBlanks, aes(x = `DATETIME`, y=`SE2int`))+
  geom_point(aes(color = MATERIAL))+
  geom_smooth(data = NoBlanks[NoBlanks$MATERIAL=='STD',], method = 'loess')+
  geom_vline(xintercept = MassCalibTime, colour = "red")+
  scale_x_datetime()+
  ylab("Measurement 2*SE (\u2030)")+
  theme_bw()+
  theme(legend.position="bottom")

d18OvsYield <- ggplot(data = NoBlanks, aes(x = `d18Omeas`, y=`Yield`))+
  geom_point(aes(color = MATERIAL))+
  xlab("\u03B418O Measured (\u2030)")+
  theme_bw()+
  theme(legend.position="bottom")

d18OvsBeam <- ggplot(data = Standards, aes(x=`IPnA`, y=d18Omeas))+
  geom_point()+
  xlab("IP (nA)")+
  ylab("\u03B418O Measured (\u2030)")+
  theme_bw()+
  theme(legend.position="bottom")

DTFAs <- ggplot(data = NoBlanks, aes(x = DTFAX, y = DTFAY, colour = MATERIAL))+
  geom_point()+
  xlab("DTFA-X")+
  ylab("DTFA-Y")+
  xlim(-50,50)+
  ylim(-50,50)+
  coord_fixed()+
  theme_bw()+
  theme(legend.position = "bottom")
  
DTFAs <- ggMarginal(DTFAs, type = "density", margins = "both", fill = "grey")

XDTFAX <- ggplot(data = NoBlanks, aes(x = X, y = DTFAX, colour = MATERIAL))+
  geom_point()+
  xlab("X \u03BCm")+
  ylab("DTFA-X")+
  theme_bw()+
  theme(legend.position="bottom")

YDTFAY <- ggplot(data = NoBlanks, aes(x = Y, y = DTFAY, colour = MATERIAL))+
  geom_point()+
  xlab("Y \u03BCm")+
  ylab("DTFA-Y")+
  theme_bw()+
  theme(legend.position="bottom")

circle <- data.frame(x.0=0,y.0=0,r=8000)

Geometery <- ggplot(data = NoBlanks, aes(x = X, y = Y, colour = MATERIAL))+
  facet_wrap(~GUESS.SAMP)+
  geom_point()+
  geom_circle(data = circle, aes(x0 = x.0, y0 = y.0, r = r), inherit.aes = FALSE)+
  xlab("X \u03BCm")+
  ylab("Y \u03BCm")+
  coord_fixed()+
  theme_bw()+
  theme(legend.position="bottom")

#### Summary statistic calculation for a variety of columns ####
#### focused on differentiating sample and standard as
#### initially identified by GREP in inport data function

Materiald18O <- summaryBy(d18Omeas~MATERIAL*GUESS.SAMP, data = NoBlanks, FUN = c(mean,sd,min,max,length))

MaterialYield <- summaryBy(Yield~MATERIAL, data = NoBlanks, FUN = c(mean,sd,min,max))

MaterialHyd <- summaryBy(Hyd~MATERIAL, data = NoBlanks, FUN = c(mean,sd,min,max))

MaterialDTFAX <- summaryBy(DTFAX~GUESS.SAMP, data = NoBlanks, FUN = c(mean,sd,min,max,length))

MaterialDTFAY <- summaryBy(DTFAY~GUESS.SAMP, data = NoBlanks, FUN = c(mean,sd,min,max,length))

MaterialX <- summaryBy(X~MATERIAL, data = NoBlanks, FUN = c(mean,sd,min,max))

MaterialY <- summaryBy(Y~MATERIAL, data = NoBlanks, FUN = c(mean,sd,min,max))
```

## Output overview

This is an autogenerated session summary Rmarkdown document for analyses on the Cameca-IMS 1280 in the [WiscSIMS laboratory](http://www.geology.wisc.edu/~wiscsims/) at the University of Wisconsin-Madison.

This document provides summary statistics on instrument stability and standard analyses within a $\delta^{18}$O session. This does not provide assessments of data quality for any single analysis.

## Summary for SIMS Session: ``r filename``

This session lasted **`r SessionTime`** hours with **`r nrow(NoBlanks)`** analyses comprising **`r nrow(Standards)`** standard analyses and **`r nrow(Samples)`** sample analyses.

The included figures and tables summarize standard and sample analyses for the session. Mount identification is infered from repeated characters in the *comment* column within the Excel table produced with the session and are not currently input by users. The number of mounts is guessed based on time between analyses and the presense of a run of 8 standards separated by a space which often indicates sample change.

Standard corrected $\delta^{18}$O, OHO, and Relative Yields values shown in these figures are recalculated based on the identification of brackets and includes *all standards* in brackets.

## Important Summary figures for lab use
\newline
```{r Fig1, echo=FALSE, warning=FALSE}

Geometery

```
\newline
**Figure 1. Spatial distribution of analysis points.** Reference circle is the extent of analyzable area in the large holder mounts [@peres_new_2013]. If using small holder mounts, use a 10 mm circle. Analyses outside of this area are potentially not reliable.
\newline
```{r Fig2, echo=FALSE, fig.height=4, fig.width=4, warning=FALSE}

Standard2SDHist

```
\newline
**Figure 2. Histogram of bracket 2SD for all standard brackets** These are from the 'Bracket Average and 2SD' rows in the Excel table. Calculation is based on the repeated analysis of the running standard.
\newline
```{r Fig3, echo=FALSE, fig.height=4, fig.width=4, warning=FALSE}

StandardMeasVio

```
\newline
**Figure 3. Histogram of average measured $\delta^{18}O$ in bracket standard lines.** These are from the 'Bracket Average and 2SD' rows in the Excel table.

```{r Fig4, echo=FALSE, fig.height=4, fig.width=4, warning=FALSE}

PrimaryBeamHist

```
\newline
**Figure 4. Histogram of primary beam stability through all analyses.** These are from all of the IP(nA) entries in the Excel File.

```{r Fig5, echo=FALSE, fig.height=4, fig.width=6.5, warning=FALSE}

PrimaryBeamTime

```
\newline
**Figure 5. Primary beam through time for full session.** These are recorded with each analysis. Red vertical lines are approximate times for mass calibration routine.

```{r d18OTime, echo=FALSE, fig.height=4, fig.width=6.5, warning=FALSE}

d18OTime

```
\newline
**Figure 5. Primary beam through time for full session.**

```{r YieldTime, echo=FALSE, fig.height=4, fig.width=6.5, warning=FALSE}

YieldTime

```
\newline
**Figure 5. Primary beam through time for full session.**

```{r SEintTime, echo=FALSE, fig.height=4, fig.width=6.5, warning=FALSE}

SEintTime

```
\newline
**Figure 5. Primary beam through time for full session.**

```{r OHOd18Ocross, echo=FALSE, fig.height=4, fig.width=4, warning=FALSE}

RelHydd18O

```
\newline
**Figure 6. Primary beam through time for full session.**

```{r RelYieldd18Ocross, echo=FALSE, fig.height=4, fig.width=4, warning=FALSE}

RelYieldd18O

```
\newline
**Figure 6. Primary beam through time for full session.**

```{r RelYieldOHOcross, echo=FALSE, warning=FALSE}

RelYieldHyd

```
\newline
**Figure 6. Primary beam through time for full session.**

## Summary tables (user focused)
The tables below are useful for the user to assess their data and mount preparation.

## Summary figures (lab focused)

## Summary tables (lab focused)

**Table 1. Session time summary.** Start, stop, active analysis, and average analysis time based on the Date and Time columns in the excel spreadsheet. Currently includes standard block analysis at the beginning of the session. Approximate 'AnalysisTime' is derived from the sum of the difference in times that are less than 15 minutes to exclude longer breaks between analysis during mass calibration or instrumental downtime.

```{r echo=FALSE, warning=FALSE, paged.print=TRUE}

kable(TimeSummary, align = 'llccc')%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

```
\newpage
\blandscape
**Table 2. Session measured $\delta^{18}$O summary** Unstandardized average and range of values.

```{r Table 2, echo=FALSE, warning=FALSE, paged.print=TRUE}

kable(Materiald18O)%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

```
\elandscape
**Table 3. Session measured Yield summary** Unstandardized average and range of values.

```{r Table 3, echo=FALSE, warning=FALSE, paged.print=TRUE}

kable(MaterialYield)%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

```

**Table 4. Session measured OHO summary.**
  
```{r Table 4, echo=FALSE, warning=FALSE, paged.print=TRUE}

kable(MaterialHyd)%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

```

**Table 5. Session DTFA-X.**
  
```{r Table 5, echo=FALSE, warning=FALSE, paged.print=TRUE}

kable(MaterialDTFAX)%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

```

**Table 6. Session DTFA-Y.**
  
```{r Table 6, echo=FALSE, warning=FALSE, paged.print=TRUE}

kable(MaterialDTFAY)%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

```


```{r Fig6, echo=FALSE, fig.height=3.5, fig.width=3.5, warning=FALSE}

d18OvsBeam

```
\newline
**Figure 6. Measured $\delta^{18}$O vs primary beam intensity.**
  
  ```{r Fig7, echo=FALSE, fig.height=3.5, fig.width=3.5, warning=FALSE}

DTFAs

```
\newline
**Figure 7. DTFA-X vs DTFA-Y.** These should be within +- 40.

```{r Fig8, echo=FALSE, fig.height=3.5, fig.width=3.5, warning=FALSE}

YDTFAY

```
\newline
**Figure 8. Distance in Y $\mu$m vs DTFA-Y value.**
  
```{r Fig9, echo=FALSE, fig.height=3.5, fig.width=3.5, warning=FALSE}

XDTFAX

```
\newline
**Figure 9. Distance in X $\mu$m vs DTFA-X value.**
